---
title: "LAGOS Spatial Analysis"
author: "Tad Trimarco"
date: "2/22/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---



```{r setup, include=FALSE}
library(tidyverse) # Tidy packages
library(sf) #Spatial package that can read and create shapefiles 
library(mapview) #Interactive maps
library(LAGOSNE) #Lots and lots of clean lake data
library(USAboundaries) #USA states and counties
```


# LAGOS Analysis


## Loading in data


### First download and then specifically grab the locus (or site lat longs)

```{r data-read}
# #Lagos download script
#LAGOSNE::lagosne_get(dest_folder = LAGOSNE:::lagos_path())


#Load in lagos
lagos <- lagosne_load()

#Grab the lake centroid info
lake_centers <- lagos$locus

# Make an sf object 
spatial_lakes <- st_as_sf(lake_centers,coords=c('nhd_long','nhd_lat'),
                          crs=4326) %>%
  st_transform(2163)


```



### Convert to spatial data
```{r}
#Look at the column names
#names(lake_centers)

#Look at the structure
#str(lake_centers)

#View the full dataset
#View(lake_centers %>% slice(1:100))

 

#Subset for plotting
subset_spatial <- spatial_lakes %>%
  slice(1:100) 

subset_baser <- spatial_lakes[1:100,]

#Dynamic mapviewer
mapview(subset_spatial)

```


### Subset to only Minnesota

```{r}
states <- us_states()

#Plot all the states to check if they loaded
#mapview(states)
minnesota <- states %>%
  filter(name == 'Minnesota') %>%
  st_transform(2163)

#Subset lakes based on spatial position
minnesota_lakes <- spatial_lakes[minnesota,]

#Plotting the first 1000 lakes
minnesota_lakes %>%
  arrange(-lake_area_ha) %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```



# In-Class work


## 1) Show a map outline of Iowa and Illinois (similar to Minnesota map upstream)

```{r}
Iowa_Illinois <- states %>%
  filter(name == 'Iowa'| name =='Illinois') %>%
  st_transform(2163)
mapview(Iowa_Illinois)

```



## 2) Subset LAGOS data to these sites, how many sites are in Illinois and Iowa combined? How does this compare to Minnesota?

```{r}
ii_lakes <- spatial_lakes[Iowa_Illinois,]
```
There are just 16,466 lakes in Iowa and Illinois combined, as opposed to the 29,038 lakes in Minnesota.


## 3) What is the distribution of lake size in Iowa vs. Minnesota?

- Here I want to see a histogram plot with lake size on x-axis and frequency on 
y axis (check out geom_histogram)

```{r}
ggplot(data = ii_lakes) +
    geom_histogram(aes(x=lake_area_ha), binwidth = 1500)

ii_lakes %>%
  arrange(-lake_area_ha) %>%
  slice(1:1000) %>%
  ggplot() +
    geom_histogram(aes(x=lake_area_ha), binwidth = 1500)

```
The vast vast majority of lakes are small, with a lake area less than 2500ha. However, there are a few larger lakes, that make the histogram difficult to view effectively. For better visualization, I have created two histograms, the second includes only the smallest 1000 lakes by size.

## 4) Make an interactive plot of lakes in Iowa and Illinois and color them
by lake area in hectares

```{r}
ii_lakes %>%
    slice(1:1000) %>%
  mapview(.,zcol = 'lake_area_ha')
```


## 5) What other data sources might we use to understand how reservoirs and 
natural lakes vary in size in these three states? 

It would also be valuable to have average depth and maximum depth measurements to better understand how large these lakes are. It may also be valuable to estimate how many rivers flow into the lake or flow out, though this isn't a perfect approximation of size, and may convey more about hydrologic connectivity.
